Tail Indices and Scale Parameters in Financial Time Series

### 2017-02-24 papers/FX/draft.tex
Adding a paragraph about the variability of the scale parameter.

\title{Tail Indices and Scale Parameters in Financial Time Series}
\author{Thomas Valentin Mikosch, Casper de Vries, Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

@@

\subsubsection[in case gamma > 1]{in case $\gamma > 1$}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience,
let $\xi = \gamma - 1$. In this case, we assume the following
distribution function $F(\cdot)$ of the equity return:
\begin{equation*}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x < 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation*}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
Equation
\eqref{eq:xxie0} can be re-arranged as
\[
u(v) + b u(\delta v))
=
\E u(C) + b \E [u(C); C < \delta v]
\]
The right-hand-side of the above equation is a functional of the
distribution function of $X$ via eq.\eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + b \E [u(C); C < \delta v]
\]
\begin{theorem}
  Let
  \[
  G(\alpha, \beta, K, K') =
  \max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
  \]
  Let the disappointment threshold $C < \delta v$ be equivalent to $X < -q$. Assume
  \begin{equation}
    \label{eq:xxie1.1}
    r > -\ln \left\{
      1 - \left[
          {\alpha + 1 \over \alpha}
          {
            \left[
              b K^{\alpha + 1} + (K + q)^{\alpha + 1}
            \right]
            \over
            \left[
              b K^{\alpha+1} \left(1 + {q \over K} \right) + (K + q)^{\alpha + 1}
            \right]
          }
          {
            1 \over K e^{1/\alpha}
          }
      \right]^{1/\xi}
    \right\}
  \end{equation}
  Then $G(\alpha, \beta, K, K')$ is monotonically increasing with $\alpha$.
\end{theorem}

@@

        \alpha K^\alpha p \left[ (K + q)^{-\alpha - 1} b + K^{-\alpha - 1} \right]

@@

@@

        e^r + {\hat\phi \over 1 - \hat\phi}

@@

        (\alpha + 1) e^{r \xi} \left[

@@

      e^r + {\hat\phi \over 1 - \hat\phi} &<&

@@

          (K + q)^{\alpha + 1} K + b K^{\alpha + 1} (K + q)

@@

          \alpha

@@

          \alpha + 1

@@

      \right]^{-1/\xi}
      \exp \left(
        r - {1 \over \alpha \xi}
      \right) \nonumber \\

@@

@@

            b K^{\alpha+1} \left(1 + {q \over K} \right) + (K + q)^{\alpha + 1}

@@

          \exp \left(
            r \xi - {1 \over \alpha}
          \right) \over K
      \right\}^{1/\xi} \nonumber \\
      \hat\phi &<& {e^r - g \over e^r - g - 1}

@@

@@

          \exp \left(
            r \xi - {1 \over \alpha}
          \right) \over K

@@

      \right\}^{1/\xi}

@@

    Assumption \eqref{eq:xxie1.1} ensures $g < e^r - 1$. Hence the
    right-hand-side of inequality \eqref{eq:xxie7} is larger than 1.
    So \eqref{eq:xxie7} is satisfied by all $\hat\phi$. Thus
    \eqref{eq:xxie6} holds and we conclude

@@

  \item When $q < 0$, we have
    \begin{eqnarray*}
      && \mathcal U(F)(\alpha, \beta, K, K', \hat\phi) \\
      &=& \alpha p K^\alpha (1 + b) \int_{-\infty}^0
      [(1 - \hat\phi) e^r + \hat\phi e^x]^{-\xi}
      (K - x)^{-\alpha-1} dx \\
      &&
      + (1 - p) \beta K'^\beta \left[
        (1 + b) \int_0^{-q} + \int_{-q}^\infty
      \right]
      [(1 - \hat\phi) e^r + \hat\phi e^x]^{-\xi} (K' + x)^{-\beta-1} dx    
    \end{eqnarray*}
  \end{enumerate}
  So the same calculation as in the previous case carries through
  with $q = 0$. The same conclusion of
  \[
  \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
  \]
  stands.
\end{proof}
\begin{remark}
  When $q = 0$, the lower-bound \eqref{eq:xxie1.1} on the interest
  rate simplifies to
  \begin{equation}
    \label{eq:xxie8}
    K < {
      (\alpha + 1) e^{-1/\alpha}
      \over
      \alpha (1 - e^{-r})^\xi
    }
  \end{equation}
  The right-hand-side of \eqref{eq:xxie8} is a very high upper bound
  on the scale parameter $K$, as is revealed later in this paper.
\end{remark}

@@

\section{Data Analysis}
\subsection{Tail index and scale parameter}
In this section we apply the statistical tools reviewed in the
previous sections to real stock return series, denoted $X_t$
hereafter. We are interested in finding out the tail indices and the
scale parameters of the stationary distribution of $|X_t|$.

>>>>

\title{Tail Indices and Scale Parameters in Financial Time Series}
\author{Thomas Mikosch, Casper de Vries, Xie Xiaolei}
\date{\today}
\begin{document}
\maketitle

@@

\subsubsection[in case gamma > 1]{in case $\gamma > 1$}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience,
let $\xi = \gamma - 1$. In this case, we assume the following
distribution function $F(\cdot)$ of the equity return:
\begin{equation*}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x < 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation*}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
Equation
\eqref{eq:xxie0} can be re-arranged as
\[
u(v) + b u(\delta v))
=
\E u(C) + b \E [u(C); C < \delta v]
\]
The right-hand-side of the above equation is a functional of the
distribution function of $X$ via eq.\eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + b \E [u(C); C < \delta v]
\]
$C < \delta v$ implies
\[
X < -q(r, \phi, \delta v) := \ln \left( {
  \delta v - (1 - \phi) e^r
  \over
  \phi
} \right)
\]
In the rest of this section we write $q$ for $q(r, \phi, \delta v)$
for convenience. We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  -{\alpha K^\alpha \over \xi} p
  \int_{-\infty}^0
      {
       1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
      }
  {1 + b \1{x < -q} \over (K + x)^{\alpha + 1}} dx \nonumber \\
  && -{\beta K'^\beta \over \xi} (1 - p)
  \int_{0}^\infty
  {
   1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
  }
  {1 + b \1{x < -q} \over (K' + x)^{\beta + 1}} dx
  \label{eq:xxie1.0}
\end{eqnarray}

\begin{theorem}
  Let
  \[
  G(\alpha, \beta, K, K') =
  \max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
  \]
  Let the disappointment threshold $C < \delta v$ be equivalent to $X < -q$. Assume
  \begin{equation}
    \label{eq:xxie1.1}
    K < {
      \alpha e^{1/\alpha} (1 - e^{-r})^\xi
      \over
      (\alpha + 1) A
    }
  \end{equation}
  where
  \[
  A = \max_{x \geq 1} {b x + x^{\alpha + 1} \over b + x^{\alpha + 1}}
  \]
  Then $G(\alpha, \beta, K, K')$ is monotonically increasing with $\alpha$.
\end{theorem}

@@

        \alpha K^\alpha p \left[

@@

        \right]^{\xi}

@@

           {\hat\phi \over 1 - \hat\phi} e^{-r} + 1

@@

        (\alpha + 1) \left[

@@

      {\hat\phi \over 1 - \hat\phi}e^{-r} + 1
      &<&

@@

          b K^{\alpha + 1} (K + q)
          + (K + q)^{\alpha + 1} K 

@@

          1

@@

          e^{1/\alpha}

@@

     \right]^{1/\xi} \nonumber \\

@@

          b \left(1 + {q \over K}\right)
          + \left(1 + {q \over K}\right)^{\alpha + 1}

@@

          b + \left(1 + {q \over K}\right)^{\alpha + 1}
          K
          e^{1/\alpha}

@@

      \right]^{1/\xi} \nonumber \\
      \hat\phi &<& {1 - g \over 1 - g - e^{-r}}

@@

      \right]^{1/\xi}
    Inequality \eqref{eq:xxie7} is satisfied by all possible
    $\hat\phi$ if
      g &<& 1 - e^{-r}
    Observe that the expression

@@

      b \left(1 + {q \over K}\right)
      + \left(1 + {q \over K}\right)^{\alpha + 1}
      b + \left(1 + {q \over K}\right)^{\alpha + 1}

@@

@@

    is 1 and increasing at $q = 0$; as $q \to \infty$, it tends to
    1. So it must have a maximum value for
    $q \in (0, \infty)$. Let
    A = \max_{x \geq 1} {b x + x^{\alpha + 1} \over b + x^{\alpha + 1}}
    Then $g < 1 - e^{-r}$ is implied by
      {K \over e^{1/\alpha}}
      A &<& (1 - e^{-r})^\xi \\
      K &<& {
        \alpha e^{1/\alpha} (1 - e^{-r})^\xi
        (\alpha + 1) A
    The last inequality is indeed assumption \eqref{eq:xxie1.1}.
    So, under this assumption, \eqref{eq:xxie7} is satisfied by all
    $\hat\phi$ and \eqref{eq:xxie6} holds. We conclude

@@

  \item When $q < 0$, it follows from \eqref{eq:xxie1.0}:
    \begin{eqnarray*}
      && \opd{\alpha}\mathcal U(F)(\alpha, \beta, K, K', \hat\phi) \\
      &=& \opd{\alpha} \left[
        -{\alpha K^\alpha \over \xi} p
        \int_{-\infty}^0
            {
              1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
            }
            {1 + b \over (K + x)^{\alpha + 1}} dx
        \right]
    \end{eqnarray*}
  \end{enumerate}
  So the same calculation as in the previous case carries through
  with 0 substituted for $q$. The same conclusion of
  \[
  \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
  \]
  stands.
\end{proof}
Since $u(v) + b u(\delta v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$, all agents guided by this class of utility
functions will seek the largest $\alpha$ found in a stock group e.g.
a sector. As a result, one can expect all the actively traded equities
in the group to feature the same $\alpha$. In section
\ref{sec:data_analysis} we analyse real-world data of stock returns to
verify this conclusion.

In comparison to the theoretical constancy of $\alpha$, the scale and
shift parameter $K$ in the lower-tail distribution of $X$ may vary
hugely from one equity to another. This could be understood also by
computing $d\mathcal U F / dK$:

@@

\section{Data Analysis} \label{sec:data_analysis}
\subsection{Tail index and scale parameter}
In this section we apply the statistical tools reviewed in the
previous sections to real stock return series, denoted $X_t$
hereafter. We are interested in finding out the tail indices and the
scale parameters of the stationary distribution of $|X_t|$.
